A comparative risk assesment of U.S. regional active transport using the ITHIM model

### 2016-10-05 R/ITHIM.Rmd
Tweaked DALYs averted figure to be just CA

```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
upper <- 25; wd <- 10; length <- upper/wd + 1;

@@

for( muNonTravel in seq(0,upper,length.out = length) ){

@@

ITHIM.NY <- ITHIMList[["NY"]]
ITHIM.NY <- updateITHIM(ITHIM.NY, parName = "muNonTravel", 10)
foo2 <- compareModels(ITHIM.national,ITHIM.NY)
foobar2 <- data.frame(cbind(melt(foo2$daly.delta),paste0("ageClass",1:8)),State = "NY")

foobar <- rbind(foobar1,foobar2)
names(foobar) <- c("delta.DALY","Sex","Disease","ageClass","State")

@@

p <- ggplot( subset(foobar,ageClass != "ageClass1"), aes(ageClass,-delta.DALY/1e3, fill = State) )

upper <- 10

p + geom_bar(stat="identity", position = "dodge") +
theme(axis.title=element_blank()) +
facet_grid( Disease ~ ., scales = "free") +
scale_y_continuous(breaks=seq(1, upper, 1), name="DALYs Averted Nationwide (Thousands)") +

#geom_hline(yintercept=seq(1, upper, 1), col="white", lwd=1) +
scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) +
theme_tufte(base_size=12, ticks=T)
#

# + annotate("text", x = 3.5, y = 5, adj=1,  family="serif", label = c("Foo"))
# + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + facet_grid( Disease ~ State) + coord_flip() # + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=8))
```

>>>>

```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
upper <- 25; length <- 10; wd <- upper/length;

@@

for( muNonTravel in seq(0, upper, by = wd)){

@@

foobar <- foobar1
names(foobar) <- c("delta.DALY","Sex","Disease","ageClass","State")

## ITHIM.NY <- ITHIMList[["NY"]]
## ITHIM.NY <- updateITHIM(ITHIM.NY, parName = "muNonTravel", 10)
## foo2 <- compareModels(ITHIM.national,ITHIM.NY)
## foobar2 <- data.frame(cbind(melt(foo2$daly.delta),paste0("ageClass",1:8)),State = "NY")
## foobar <- rbind(foobar1,foobar2)
## names(foobar) <- c("delta.DALY","Sex","Disease","ageClass","State")

@@

p <- ggplot( subset(foobar,ageClass != "ageClass1"), aes(ageClass,-delta.DALY/1e3, fill = Sex) )

#upper <- 10

p + geom_bar(stat="identity", position = "dodge") +
theme(axis.title=element_blank()) +
facet_grid( Disease ~ ., scales = "free") +
scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") +
scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+"))

#geom_hline(yintercept=seq(1, upper, 1), col="white", lwd=1) +
# + theme_tufte(base_size=12, ticks=T)
#
# + annotate("text", x = 3.5, y = 5, adj=1,  family="serif", label = c("Foo"))
# + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + facet_grid( Disease ~ State) + coord_flip() # + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=8))
```

### 2016-10-05 manuscript/manuscript.tex
Removed commented-out bits
Tweaked DALYs averted figure to be just CA

\section{Availability:}

\package{} may be explored using the user interface avaiable
at \webpage{}.  The source code and data files for \package{} are
publicly available on GitHub at \webpage{}.

@@

Of these seven states one was found to show active transport behavior
that, if adopted nation-wide, would avert 75-100,000 DALYs overall.
Values for national disease-specific DALYs were found using the CDC
Wonder database \cite{CDCWonder}.  California stood out among the
estimates for national DALYs averted.  Most of the decrease in DALYs
is due to prevention of dementia in the oldest age group, $80+$ years
old.  California exhibits much greater walking means in this age group
than the national average.  We also see prevention of depression and
diabetes in the middle age groups due to increased active transport
time.

@@

%% \subsection{Road Injuries}
%% \textit{ We chose to model road injuries in a more simplifed manner
%%   than the authors of ITHIM.  The original ITHIM requires information
%%   about the transportational infrastructure that cannot be accurately
%%   modeled on a regional level, and is more well-suited to say a
%%   collection of counties as a case study. \cite{maizlish2013} We use
%%   the overall percent of travel by walking and road fatalities per
%%   U.S. metropolitan area to estimate road fatalaties in
%%   U.S. metropolitan areas given the percent of travel by
%%   walking. \cite{UWTechReport}}

%% \textit{ We need to estimate the increase/decrease in road
%%   injuries/fatalaties nation-wide as walking/cycling time increases.
%%   We need a relative risk of road fatalaty as a function of active
%%   transport time\ldots}

%% \subsection{Air Pollution}

%% \textit{ Can we model the decrease in \pEm{2.5} nationally as a
%%   function of decrease in driving time?  The original plan was to use
%%   a baseline value for \pEm{2.5} and a linear relationship bewteen
%%   change in \pEm{2.5} and change in \vmt{}.}
%% \begin{figure}[t]
%%     \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig5.pdf}}
%%     \caption{Counts}\label{counts}
%% \end{figure}

%% \begin{itemize}
%% \item We used the National Household Transportation Survey \cite{NHTS} to
%%   estimate the mean walk and cycle times for individuals that live in
%%   metropolitan areas ($\textrm{DAYV2PUB.MSA} \in (1,2,\ldots,5)$ which
%%   corresponds to individuals whose homes are in an MSA or cMSA)
%%   \bi\item Data quality issues along with assumptions about travel speed
%% created mirky data.  The median values of trip time were used to avoid
%% undue influence by outliers.\ei
%% \item We then divided these into the 10 regions of the U.S. defined by
%%   the Department of Health and Human Services (HH1-10)
%% \item Using only the active transport component of the ITHIM model we
%%   ran comparative analyses for each of the ten regions, using the
%%   national averages as the baseline.
%% \item We used data from the American Time Use Survey \cite{ATUS} to
%%   estimate the relative means for each of the age-sex strata.  We then
%%   treat the mean of the referent group as a variable and compute
%%   results over a grid of values for this variable.  See Figure
%%   \ref{dalyFigure}.
%% \item 
%% \end{itemize}

>>>>

%% \section{Availability:}

%% %% \package{} may be explored using the user interface avaiable
%% %% at \webpage{}.
%% The source code and data files for \package{} are publicly available
%% on GitHub via a link on \webpage{}.

@@

Of these seven states one was found to show active transport behavior
that, if adopted nation-wide, would avert 75-100,000 DALYs overall.
Values for national disease-specific DALYs were found using the CDC
Wonder database \cite{CDCWonder}.  California stood out among the
estimates for national DALYs averted.  Most of the decrease in DALYs
is due to prevention of dementia in the oldest age group, $80+$ years
old.  California exhibits much greater walking and cycling means in this age group
than the national average. In particular we see the cycling mean We also see prevention of depression and
diabetes in the middle age groups due to increased active transport
time.

@@
